---
title: "Reproducible Research"
author: "Mudit Maheshwari"
date: "September 20, 2015"
output: html_document
---

```{r, echo = FALSE, results = "hide"}
library(dplyr)
library(tidyr)
```

Loading and preprocessing the data:

```{r}
activity <- read.csv("C:/R/Coursera/05_ReproducibleResearch/activity.csv")
```

What is mean total number of steps taken per day?:

```{r, echo=TRUE}
activity_noNas <- activity[complete.cases(activity),] 
activity_summary <- activity_noNas %>% group_by(date) %>% 
                  summarise(steps = sum(steps))

summary <- activity_summary %>% summarise(Total_Steps = sum(steps),
                               Mean_Steps = mean(steps),
                               Median_Steps = median(steps)
                               )

hist(activity_summary$steps, 
     main="Histogram for Activity", 
     xlab="Steps", 
     border="blue", 
     col="green"
     )
Total_Steps <- summary$Total_Steps

```


What is the average daily activity pattern?

```{r, echo=TRUE}

activity_interval <- activity_noNas %>% group_by(interval) %>% 
                  summarise(steps = mean(steps))

plot(activity_interval$interval, activity_interval$steps, type = 'l', xlab = "Interval",
     ylab = "Steps", axes = FALSE)
     axis(side = 1, c(seq(from=0,to=2500,by=100)))
     axis(side = 2, c(seq(from=0,to=250,by=25)))
     
Interval_Max <- activity_interval %>% filter(steps >= max(steps)) %>%
                      select(interval)
          
```

The interval that contains maximum number of steps is `r Interval_Max`


Imputing missing values:

```{r, echo=TRUE}
Steps_NA_CNT <- sum(is.na(activity$steps))
```

The total number of rows with NAs are `r Steps_NA_CNT`

```{r, echo=TRUE}
activity_Nas <- activity[is.na(activity$steps), 2:3]

activity_Nas_Imputed <- activity_noNas %>% group_by(interval) %>%
                   summarise(steps = median(steps)) %>%
                   inner_join(activity_Nas, by = c("interval" = "interval")) %>%
                  select(steps, date, interval)
  
activity_imputed <- bind_rows(activity_Nas_Imputed, activity_noNas)

activity_summary_Imputed <- activity_imputed %>% group_by(date) %>% 
                  summarise(steps = sum(steps))

summary_imputed <- activity_summary_Imputed %>% summarise(Total_Steps = sum(steps),
                               Mean_Steps = mean(steps),
                               Median_Steps = median(steps)
                               )

hist(activity_summary_Imputed$steps, 
     main="Histogram for Activity", 
     xlab="Steps", 
     border="blue", 
     col="green"
     )

Total_Steps_Imputed <- summary_imputed$Total_Steps 

```


The total steps after imputing for NAs were `r Total_Steps_Imputed` where as total steps before imputing were `r Total_Steps`


Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}

activity_imputed$date <- as.Date(activity_imputed$date)
#create a vector of weekdays
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
#convert to `factor` and specify the `levels/labels`
activity_imputed$wDay <- factor((weekdays(activity_imputed$date) %in% weekdays1), 
         levels=c(FALSE, TRUE), labels=c('weekend', 'weekday') )


activity_interval_wday <- activity_imputed %>% group_by(wDay, interval) %>% 
                  summarise(steps = mean(steps))

library(lattice)
xyplot(steps ~ interval | wDay, activity_interval_wday, layout = c(2,1), type = 'l')

```
